### 2014-01-26 tex/gecco14/gecco14.tex
gecco atd

\documentclass{sig-alternate}

\usepackage[utf8]{inputenc}
\usepackage{xspace}
\usepackage[ampersand]{easylist}

\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}

\newenvironment{lizt}
{\begin{easylist}[itemize]}
{\end{easylist}}

\newcommand{\Lets}{Let us\xspace}
\newcommand{\lets}{let us\xspace}
\newcommand{\lterm}{$\lambda$-term\xspace}
\newcommand{\lterms}{$\lambda$-terms\xspace}
\newcommand{\lhead}{$\lambda$-head\xspace}
\newcommand{\lheads}{$\lambda$-heads\xspace}
\newcommand{\la}{\leftarrow\xspace}
\newcommand{\Lp}  {\Lambda^{\prime}\xspace}
\newcommand{\tur}[3]{#1\vdash{}#2 \colon #3}
\newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
\newcommand{\atTree}{@-tree\xspace}
\newcommand{\setDots}[2]{ \lbrace #1 , \dots , #2 \rbrace}
\newcommand{\lh}[1]{\lambda #1}
\newcommand{\sexprTree}{sexpr-tree\xspace}
\newcommand{\SexprTree}{Sexpr-tree\xspace}
\newcommand{\then}{\Rightarrow\xspace}
\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
\newcommand{\ST}{\mathop{\mathrm{ST}}}
\newcommand{\FV}{\mathop{\mathrm{FV}}}
\newcommand{\Scomb }{\mathbf{S}}
\newcommand{\Kcomb }{\mathbf{K}}
\newcommand{\Icomb }{\mathbf{I}}
\newcommand{\bbarr}{\twoheadrightarrow_\beta}
\newcommand{\barr}{\rightarrow_\beta}
\newcommand{\beq}{=_\beta}
\newcommand{\eearr}{\twoheadrightarrow_\eta}
\newcommand{\earr}{\rightarrow_\eta}
\newcommand{\eeq}{=_\eta}
\newcommand{\bearr}{\rightarrow_{\beta\eta}}
\newcommand{\bbeearr}{\twoheadrightarrow_{\beta\eta}}
\newcommand{\beeq}{=_{\beta\eta}}
\newcommand{\etar}{\twoheadrightarrow_\eta}
\newcommand{\ered}{$\eta$-reduction\xspace}
\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
\newcommand{\bredex}{$\beta$-redex\xspace} 
\newcommand{\lnf}{\textit{lnf}\xspace}
\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
\newcommand{\Bcomb }{\mathbf{B}}   
\newcommand{\BBcomb }{\mathbf{B*}}
\newcommand{\Ccomb }{\mathbf{C}}   
\newcommand{\CCcomb }{\mathbf{C'}}
\newcommand{\SScomb }{\mathbf{S'}}
\newcommand{\ar}{\rightarrow\xspace}
\newcommand{\T}{\mathbb{T}\xspace}
\newcommand{\C}{\mathbb{C}\xspace}
\newcommand{\Real}{\mathbb{R}}
\newcommand{\gar}{\longmapsto}

\newenvironment{todo}
{~\\ {\color{red}\textbf{TODO}}
  \begin{easylist}[itemize]}
{ \end{easylist}}

\newcommand{\Lpr}{\Lambda^\prime}
\newcommand{\ul}[2]{\langle #1 ; #2 \rangle}
\newcommand{\ro}[1]{{\color{blue} #1}}
\newcommand{\tom}[1]{{\color{ForestGreen} #1}}
\newcommand{\red}[1]{{\color{red} #1}}

\begin{document}

\conferenceinfo{GECCO'14,} {July 12-16, 2014, Vancouver, BC, Canada.}
\CopyrightYear{2014}
\crdata{TBA}
\clubpenalty=10000
\widowpenalty = 10000

\title{Utilization of Reductions and Abstraction Elimination in Typed Genetic Programming}

\numberofauthors{2}

\author{
% 1st. author
\alignauthor
Tom\'{a}\v{s} K\v{r}en\\
  \affaddr{Faculty of Mathematics and Physics}\\
  \affaddr{Charles University in Prague}\\
  \affaddr{Malostransk\'e n\'am\v{e}st\'\i~25, 11000,}\\
  \affaddr{Prague, Czech Republic}\\
  \email{tomkren@gmail.com}
% 2nd. author
\alignauthor
Roman Neruda\\
  \affaddr{Institute of Computer Science}\\
  \affaddr{Academy of Sciences of the Czech Republic}\\
  \affaddr{Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207,}\\
  \affaddr{Prague, Czech Republic}\\
  \email{roman@cs.cas.cz}
}

%\date{24 January 2014}

\maketitle
\begin{abstract}
Lambda calculus representation of programs offers a more expressive alternative to traditional S-expressions. In this paper we discuss advantages of this representation coming from use of reductions (beta and eta) and how to overcome disadvantages caused by variables occurring in the programs by use of the abstraction elimination algorithm. We discuss the role of those reductions in the process of generating initial population and compare several crossover approaches including novel approach to crossover operator based both on reductions and abstraction elimination. The design goal of this operator is to turn the disadvantage of abstraction elimination  - possibly quadratic increase of program size - into a virtue; our approach leads to more crossover points. At the same time, utilization of reductions provides offspring of small sizes.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{TODO}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{TODO}

\keywords{TODO, ToDo, todo}

\section{Plán boje}

\subsection{Jak to celý pojmout?}

Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama. Motivace pro práci s lambda termama je, že za nima košatá teorie, která například popisuje redukce lambda termů (ty zajišťujou jak zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, díky tomu, že se různé stromy redukují na ten samej).

Jak se vypořádat s proměnejma a lambdama? (aka jak křížit)

\begin{itemize}
\item Po vy generování převod do kombinátorů pomocí eliminace abstrakcí (tak jak se to dělá v diplomce). To ale působí v něčem neohrabaně, když to porovnáme s generovánim přímo v kombinatorech, když vezmeme v potaz to že eliminace abstrakcí má za důsledek až kvadratickej nárůst stromu - čili to co sme nahnali na redukcích stratíme eliminací.

\item Jedince držíme jako redukovane malinké-kompaktní-a-elegantní lambda termy. Ve chvíly kdy křížíme provedeme na obou rodičích eliminaci abstrakcí (která ubere proměny a lambdy a namísto toho tam dá kombinátory S,K,I případně i další při fikanějších eliminacích), tím nám sice narostou, ale my z toho máme jedine radost, protože tím se nám zvýšil počet míst ke křížení (což se ukazuje jako dobráv věc, viz s-expr reprezentace vs @-tree reprezentace lambda termů). Po skřížení se vložený kombinátory nahradí odpovídajícim lambda termem (tzn např všude kde je K dám (\ x y . x) atd) výslednej term zredukuju a dostávam zase malinké-kompaktní-a-elegantní dítě.
Nevýhoda toho zahrnout do článku i tohle je v tom, že k tomu nemám ještě žádný pokusy - ale k tomu zbytku mám upřímě v zato taky dost ubohý pokysy, takže toho bych se asi nebál. Většinu potřebnýho kodu bych k tomu ale měl už mít víceméně hotovou, takže pokud by se to na něčem nezaseklo, tak myslim že je realný udělat i pokus do toho dvacátýho. Zvlášť přitažlivý mi tohle křížení příde i kvůli tomu, že v přírodě se taky rozbalujou a zabaloujou chromozomi při meioze/mitoze.
\end{itemize}

\subsection{Osnova}

\begin{lizt}
& Jak reprezentovat stromy programu pro GP?
  && Reprezentace v klasickym kozovy je S-expression.
  && Nebo mužem používat kombinátory jako 
     Briggs a O’Neil (to si myslim je jakoby 
     hlavní konkurence, vuči který by to chtělo obhájit)
  && Lambda termy a jejich awesomeness
& \textbf{Povídaní o redukcích}
& Povídání o lnf
  && že lnf je přirozený rozšíření s-exprešnu do lambda 
     kalkulu vlastnosti termu v lnf 
     &&& proč je eta redukovat (eta redukcí lnf dostanem 
         beta-eta-nf a že nemusíme beta redukovat pač se to tim nerozbyje)
& Generování
  && Generování gramatikou
  && Gramatika pro lnf
  && (?) Inhabitation trees jako intuitivní model takovýhodle lnf generování
& Problémy s proměnýma a jak je řešit.
  && Křížit lambda termy i s proměnejma a abstrakcema, 
     problémy řešit když nastanou (pomluvit a odsoudit)
  && Zmenšit prostor termu (tim že někerý nejsme schopný 
     vygenerovat) s kterým operujeme tak aby křížení už 
     nebyl problém (to dělá Yu - v těle lambda fce dovoluje 
     jen použití proměnných z její hlavy)
  && Převod hned po vygenerování
  && Převod až při křížení
     &&& eliminace abstrakcí
     &&& skřížim
     &&& vložený kombinátory nahradim odpovídajícím termem
     &&& celý to redukuju
 & Pokusy (?!)
 & Závěr
\end{lizt}

\section{Introduction}

Genetic programming (GP) represents an efficient method for automatic generating of programs by means of evolutionary techniques~\cite{koza92,koza03}. Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.   
Use of types naturally opens door to enriching S-expressions,
the traditional GP representation of individuals, with concepts from
lambda calculus, which is simple yet powerful functional mathematical and programming 
language extensively used in type theory. Such attempts has shown to be 
successful \cite{yu01}.

The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation. During the expansion phase the set of unfinished terms can be browsed with respect to various search strategies. Our approach to this problem aims to utilize the full arsenal given by the simply typed lambda calculus. Thus, the natural idea is to employ an exhaustive systematic search. On the other hand, if we were to mimic the standard GP approach, a quite arbitrary yet common and successful ramped half-and-half generating heuristic~\cite{fg} should probably be used.  
These two search methods in fact represent boundaries between which we will try to position our parameterized solution that allows us to take advantage of both strategies. This design goal also differentiate our approach from 
the three state of the art proposals for typed GP known to us that are discussed in the following section. 
Our proposed \emph{geometrical search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search. Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.

The rest of the paper is organized as follows: The next section briefly discusses related work in the field of typed GP, while section~\ref{preliminaries} introduces necessary notions. Main original results about search strategies in individual generating are described in section~\ref{approach}. Section~\ref{experiments} presents results of our method on three well-known tasks, and the paper is concluded by section~\ref{conclusions}.

\section{Related work}
\label{related}

Yu presents a GP system utilizing
polymorphic higher-order functions\footnote{Higher-order 
function is a function taking another function as 
input parameter.} and lambda abstractions  \cite{yu01}.
Important point of interest in this work is use of
\texttt{foldr} function as a tool for \textit{implicit recursion},
i.e. recursion without explicit recursive calls. 
The terminal set for constructing lambda abstraction subtrees 
is limited to use only constants and variables of that particular
lambda abstraction, i.e., outer variables are not allowed to be used
as terminals in this work. This is significant difference from our approach 
since we permit all well-typed normalized \lterms. From this difference also
comes different crossover operation. We focus more on term generating process; 
their term generation is performed in a similar way as the standard one, 
whereas our term generation also tries to utilize techniques of systematic enumeration.

Briggs and O’Neill present technique 
utilizing typed GP with combinators \cite{kes}.
The difference between approach presented in this work
and our approach is that in this work terms are generated
straight from \textit{library} of combinators and no lambda abstractions
are used. They are using more general polymorphic type system than us
-- the Hindley–Milner type system. They also discuss the 
properties of exhaustive enumeration of terms and compare it with GP search.  
They also present interesting concept of \textit{Generalized
genetic operator} based on term generation.

Binard and Felty use even 
stronger type system (\textit{System F}) \cite{binard2008genetic}.  
But with increasing power of the type system comes increasing difficulty of term generation.
For this reason evolution in this work takes interesting and nonstandard shape 
(fitness is associated with \textit{genes} which are evolved together with \textit{species}
which together participate in creation of individuals).
This differs from our approach, which tries to be generalization of
the standard GP\cite{koza92}.

In contrast with above mentioned works our approach uses very simple type system 
(simply typed lambda calculus) and concentrates on process of generation  
able to generate all possible well-typed normalized lambda terms. In order to do
so we use technique based on \textit{inhabitation machines} 
described by Barendregt \cite{barendregt10}.

\section{Preliminaries}
\label{preliminaries}

In this section, several notions necessary to build a typed GP based on lambda calculus are introduced. 
First, \lets describe a programming language, 
in which the GP algorithm generates individual programs --- the so called \lterms.

\begin{definition}
Let $V$ be infinite countable set of {\it 
variable names}. Let $C$ be set of {\it constant names}, 
$V \cap C = \emptyset$.	 	
Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
\begin{align*}
x   \in V \cup C  &\then x     \in \Lambda \\
M,N \in \Lambda   &\then (M~N) \in \Lambda 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
\textit{~~~~($\lambda$-abstraction)} 
\end{align*}
\end{definition}

\textit{Function application} and 
\textit{$\lambda$-abstraction} are concepts
well known from common programming languages. 
For example in JavaScript 
$(M~N)$ translates to expression \texttt{$M$($N$)} and
$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
In other words, the function application 
corresponds to the act of supplying a function 
with an argument, and
the $\lambda$-abstraction is equivalent to 
\textit{anonymous function}\footnote{Apart from JavaScript, anonymous functions are common e.g. in Python and Ruby, 
they were recently introduced to C++, and they are expected to be supported in Java 8.}.

For better readability, 
$M_1~M_2~M_3~\dots~M_n$ is an abbreviation for
$(\dots((M_1~M_2)~M_3)~\dots~M_n)$
and $\lam{x_1 x_2 \dots x_n }{M}$ for 
$\lamb{x_1}{\lamb{x_2}{\dots\lamb{x_n}{M}\dots}}$.

\subsection{$\beta$-reduction}

In order to perform computation there must be some
mechanism for term evaluation. In $\lambda$-calculus there
is $\beta$-reduction for this reason.\\

\newcommand{\bRedex}{$\beta$-redex\xspace}
\newcommand{\bRedexes}{$\beta$-redexes\xspace}
\newcommand{\bArrow}{\rightarrow_\beta\xspace}
\newcommand{\eArrow}{\rightarrow_\eta\xspace}
\newcommand{\eeArrow}{\rightarrow_{\eta^{-1}}\xspace}

A term of a form $\lamb{x}{M}N$ is called \textit{\bRedex}.
A \bRedex can be $\beta$-reduced to term $M[x:=N]$. 
This fact is written as \textit{relation} $\bArrow$ 
of those two terms:
\begin{equation} \label{eq:bRed}
\lamb{x}{M}N \bArrow M[x:=N]
\end{equation}
It is also possible to reduce \textit{subterm \bRedexes} 
which can be formally stated as:
\begin{align*}
P \bArrow Q &\then (R~P)      \bArrow (R~Q) \\
P \bArrow Q &\then (P~R)      \bArrow (Q~R) \\
P \bArrow Q &\then \lam{x}{P} \bArrow \lam{x}{Q}  
\end{align*}

In other words, $\beta$-reduction is the process 
of insertion of arguments supplied to a function into 
its body. \\

Another useful relations are $\bbarr$ and $\beq$ defined as follows.

\begin{enumerate}
 \item \begin{enumerate}
 	\item $M \bbarr M$
 	\item $M \barr N \then M \bbarr N$
 	\item $M \bbarr N , N \bbarr L \then M \bbarr L$ 	
 \end{enumerate}
 \item \begin{enumerate}
 	\item $M \bbarr N \then M \beq N$
 	\item $M \beq N \then N \beq M$
 	\item $M \beq N , N \beq L \then M \beq L$
 \end{enumerate}

\end{enumerate}

We read those relations as follows.
\begin{enumerate}
 	\item $M \bbarr N$ --- "$M$ $\beta$-reduces to $N$."  
 	\item $M \barr N$  --- "$M$ $\beta$-reduces to $N$
 	      in one step."
 	\item $M \beq N$ --- "$M$ is $\beta$-convertible to $N$."	
 \end{enumerate}

\subsection{$\eta$-reduction}

Similarly as for $\beta$-reduction we can define $\eta$-reduction 
except that instead of \ref{eq:bRed} we use:

$$\lamb{x}{(M~x)} \eArrow M \textbf{ ~~~~if } x \not\in FV(M) $$~

Analogically, a term of a form $\lamb{x}{(M~x)}$ is called 
\textit{$\eta$-redex}.

Relation $\bearr\;=\;\barr \cup \earr$. 
(Relation $R = \{\;(a,b)\;|\;a\;R\;b\;\}$.)

Similarly as for $\bbarr$ and $\beq$ we can define relations 
$\eearr$, $\eeq$,$\bbeearr$ and $\beeq$.

\subsection{$\eta^{-1}$-reduction}

$\eta^{-1}$-reduction (also called $\eta$-expansion) is 
the reduction converse to $\eta$-reduction.
Again it may be obtained by replacing \ref{eq:bRed}, now with:

$$M \eeArrow \lamb{x}{(M~x)} \textbf{ ~~~~if } x \not\in FV(M) $$

\subsection{Normal forms}

~\begin{easylist}[enumerate]
& A \lterm is a \textit{$\beta$-normal form} (\bnf) 
if it does not have a $\beta$-redex as subterm.
& A \lterm M \textit{has} a \bnf if $M \beq N$
and $N$ is a \bnf.\\
\end{easylist}
A normal form may be thought of as a result of a term evaluation.

Similarly we can define \enf and \benf.

\subsection{Types etc}

A \lterm as described above
corresponds to a program expression with no type information
included. Now we will describe \textit{types} (or \textit{type terms}).

\begin{definition}
Let $A$ be set of {\it atomic type names}. 
Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
\begin{align*}
\alpha      \in A  &\then   \alpha \in \T \\
\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
\end{align*}
\end{definition}

Type $\sigma \ar \tau$ is type for functions taking as input
something of a type $\sigma$ and returning 
as output something of a type $\tau$. 
$\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ is an abbreviation for 
$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$.
The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
combining the previously defined \textit{\lterms} and \textit{types} together.

\begin{definition}~

\begin{enumerate}
 \item 	Let $\Lambda$ be set of {\it \lterms}. 
	Let $\mathbb{T}$ be set of {\it types}.       
	A {\it statement} $M : \sigma$ is a pair 
	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
	Statement $M : \sigma$ is vocalized as 
	{\it "$M$ has type $\sigma$"}.
	The term $M$ is called the {\it subject} of the 
	statement $M : \sigma$.
 \item A \textit{declaration} is a statement 
 $x : \sigma$ where $x \in V \cup C$.

 \item A \textit{context} 
 is set of declarations with distinct variables as subjects.
\end{enumerate}
\end{definition}
%~\\

Context is a basic type theoretic concept suitable as a typed alternative
for terminal and function set in standard GP. 
Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
such that $\Gamma$ does not contain any declaration with $x$ as subject.
We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.

\begin{definition}
A statement $M\colon\sigma$ is \textit{derivable from}
a context $\Gamma$ (notation 
\mbox{$\Gamma\vdash{}M\colon\sigma$}) 
if it can be produced by the following rules.
\begin{align*}
x : \sigma \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
\tur{\Gamma,x:\sigma}{M}{\tau}
&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
\end{align*}
\end{definition}

Přeformulovat aby se to hodilo do tohodle kontextu....
Our goal in term generation is to produce terms $M$
for a given pair $\ul{\tau}{\Gamma}$
such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.

\subsection{Long normal form}
\label{lnf}

%v barendrechtovi je to v definici lnf
%$\lam{x_1 \dots x_n}{f~M_1~\dots~M_n}$ místo $M_m$ 
%ale to musí bejt určitě překlep...

\begin{definition}
Let \GMS where 
$\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha, n \geq 0$.
	\begin{enumerate}
	  \item	
		Then $M$ is in \textit{long normal form} (\lnf) if following 
		conditions are satisfied.
		\begin{enumerate}
		 \item $M$ is term of the form $\lam{x_1 \dots x_n}{f~M_1~\dots~M_m}$\\
		  (specially for $n = 0$, $M$ is term of the form $f$).
		 \item Each $M_i$ is in \lnf.
		\end{enumerate}	
	  \item 
	    $M$ has a \lnf if $M =_{\beta\eta} N$ and $N$ is in \lnf.
	\end{enumerate}
\end{definition}~

As is shown in \cite{barendregt10}, \lnf has following nice properties.

\begin{proposition}
If $M$ has a \bnf, 
%which according to Theorem 2B.4 is always the case, 
then it also has a unique \lnf, 
which is also its unique \beenf.
\end{proposition}

\begin{proposition}
Every $B$ in \bnf has a \lnf 
$L$ such that $L \twoheadrightarrow_{\eta} B$.
\end{proposition}

\subsection{Grammar producing \lterms in \lnf}

In \cite{barendregt10} is shown term generating grammar with 
following rules (our notation is used, but we will not 
highlighted terminals anymore).

%\footnote{
%I was using term generating grammars for term generation
%on my own before I encountered with \cite{barendregt10},
%where i happily discovered that Barendregt is using
%almost identical notation. 
%But his \lnf grammar was more clever then my grammar
%that was using grammar rules corresponding to the three basic
%inference rules. 
%}

\begin{align*}
( \alpha , \Gamma )  
&\gar
(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
& \textbf{if } \alpha \in A,
(f : \rho_1 \ar \dots \ar \rho_m \ar \alpha) \in \Gamma
\\ 
( \sigma \rightarrow \tau , \Gamma )  
&\gar
(~\lambda~x~.~( \tau ; \Gamma,x:\sigma )~)
&   
\end{align*}

The second rule can be replaced by more effective one.
\[ 
( \tau_1 \ar \dots \ar \tau_n \ar \alpha , \Gamma )  
\gar
(~\lambda~x_1~\dots~x_n~.~
( \alpha ; \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n  )~)
~~~~ \textbf{if } n > 0
\]

This rule packs consecutive uses of the second rule into one use.
This is valid since the use of the second rule is deterministic;
it is used if and only if the non-terminal's type is not atomic.

\section{Our approach}
\label{approach}

\section{Experiments}
\label{experiments}

\section{Conclusions}
\label{conclusions}

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{evogp}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%

% That's all folks!
\end{document}